Simple repression draft

### 2020-05-29 doc/section_04_bayesian_inference.tex
RP carefully through inference
did repressor section

We also note that we found it necessary to exclude the very
weakly and very strongly repressed datasets from Jones et.\
al.~\cite{Jones2014}. In both cases there was, in a sense, not
enough information in the distriubtions for our inference
algorithm to extract, and their inclusion simply caused problems
for the MCMC sampler without yielding any new insight. For the
strongly repressed data (Oid, 10~ng/mL aTc), with >95\% of cells
with zero mRNA, there was quite literally very little data from
which to infer rates. And the weakly repressed data, all with the
repressor binding site O3, had an unbinding rate so fast that the
sampler essentially sampled from the prior; the likelihood had
negligible influence, meaning the data was not informing the
sampler in any meaninful way, so no inference was possible.

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons,
with the top panel comparing to equilbrium binding energies
and the bottom panel comparing to single-molecule measurements.

First consider the top panel and its comparison between repressor
kinetic rates and binding energies.
As described in section~\ref{section_02_means}, if the
equilibrium binding energies from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018} indeed are the physically real binding
energies we believe them to be, then they should be related to
the repressor kinetic rates via
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we reprint here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action
kinetics implies that $k_R^+$ is proportional to repressor copy
number $R$, or more precisely, it can be thought of as repressor
copy number times some intrinsic per molecule association rate.
But since $R$ is not directly known for our data
from~\cite{Jones2014}, we cannot use this equation directly.
Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them,
since this difference depends only on the unbinding rates and not
on the binding rates. This can be seen by evaluating
Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for two different
repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal
and vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}.
Since we inferred rates for three repressor binding sites (O1, O2, and Oid),
there are only two independent differences that can be constructed, and we
arbitrarily chose to plot O2-O1 and O1-Oid in
Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and
$k_{O2}^- / k_{O1}^-$ directly from our full posterior samples,
which conveniently provides uncertainties as well, as detailed in
Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c
sampling makes marginalization trivial.}
We then compare these ratios of rates to the binding energy differences
$\Delta\epsilon_{O1} - \Delta\epsilon_{Oid}$ and from
$\Delta\epsilon_{O2} - \Delta\epsilon_{O1}$ as computed from the values
from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}.
Three of the four values are within $\sim0.5 k_BT$ of the diagonal
representing perfect agreement, which is comparable to the
$\sim\text{few}\times0.1 k_BT$ variability between the independent
determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against
high background since Oid represses so strongly. We are therefore
inclined to regard the failure of this point to fall near the
diagonal as a testament to the difficulty of the measurement and
not as a failure of our theory.
\mmnote{Does this seem a fair characterization?}

On the whole then, we regard this as striking confirmation of the
validity of the equilibrium models. Their lynchpin parameter is a
phenomenological free energy of repressor binding that has
previously only been inferred indirectly. Our result shows that
the microscopic interpretation of this free energy, as the log of
a ratio of transition rates, does indeed hold true.
\mmnote{Does this deserve qualifiers, like ``to a good approximation,'' or ``as a first order model'' or something? I like the simple finality without, but I worry it's too brazen or subtlely misleading.}

>>>>

We also note that we found it necessary to exclude the very
weakly and very strongly repressed datasets from Jones et.\
al.~\cite{Jones2014}. In both cases there was, in a sense, not
enough information in the distriubtions for our inference
algorithm to extract, and their inclusion simply caused problems
for the MCMC sampler without yielding any new insight. For the
strongly repressed data (Oid, 10~ng/mL aTc), with $>$ 95\% of cells
with zero mRNA, there was quite literally very little data from
which to infer rates. And the weakly repressed data, all with the
repressor binding site O3, had an unbinding rate so fast that the
sampler essentially sampled from the prior; the likelihood had
negligible influence, meaning the data was not informing the
sampler in any meaninful way, so no inference was possible.

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons,
with the top panel comparing to equilbrium binding energies
and the bottom panel comparing to single-molecule measurements.
First consider the top panel and its comparison between repressor
kinetic rates and binding energies.
As described in section~\ref{section_02_means}, if the
equilibrium binding energies from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018} indeed are the physically real binding
energies we believe them to be, then they should be related to
the repressor kinetic rates via
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action
kinetics implies that $k_R^+$ is proportional to repressor copy
number $R$, or more precisely, it can be thought of as repressor
copy number times some intrinsic per molecule association rate.
But since $R$ is not directly known for our data
from~\cite{Jones2014}, we cannot use this equation directly.
Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them,
since this difference depends only on the unbinding rates and not
on the binding rates. This can be seen by evaluating
Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for two different
repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal
and vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}.
Since we inferred rates for three repressor binding sites (O1, O2, and Oid),
there are only two independent differences that can be constructed, and we
arbitrarily chose to plot O2-O1 and O1-Oid in
Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and
$k_{O2}^- / k_{O1}^-$ directly from our full posterior samples,
which conveniently provides uncertainties as well, as detailed in
Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c
sampling makes marginalization trivial.}
We then compare these ratios of rates to the binding energy differences
$\Delta\epsilon_{O1} - \Delta\epsilon_{Oid}$ and from
$\Delta\epsilon_{O2} - \Delta\epsilon_{O1}$ as computed from the values
from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}.
Three of the four values are within $\sim0.5 k_BT$ of the diagonal
representing perfect agreement, which is comparable to the
$\sim\text{few}\times0.1 k_BT$ variability between the independent
determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against
high background since Oid represses so strongly. We are therefore
inclined to regard the failure of this point to fall near the
diagonal as a testament to the difficulty of the measurement and
not as a failure of our theory.
\mmnote{Does this seem a fair characterization?   - RP: thanks, that was very helpful
for me.}

On the whole then, we regard this as striking confirmation of the
validity of the equilibrium models. Their lynchpin parameter is a
phenomenological free energy of repressor binding that has
previously only been inferred indirectly. Our result shows that
the microscopic interpretation of this free energy, as the log of
a ratio of transition rates, does indeed hold true to within the inherent
uncertainties that remain in the entire theory-experiment dialogue.
%\mmnote{Does this deserve qualifiers, like ``to a good approximation,'' or ``as a first order model'' or something? I like the simple finality without, but I worry it's too brazen or subtlely misleading.}
